Publisher League Global

### 2019-09-08 analysis/004_kb_fetch_publisher.Rmd
initial upload

---
title: "Fetching Reprint Author info - v2"
output: github_document
---

@@

Country information about corresponding authors play a crucial role in open access funding (Schimmer, Geschuhn, and Vogler 2015). The Web of Science contains data about this author role including affiliations, making it a valuable source to determine publisher's market shares per country in terms of articles published.

In this work, a dataset is compiled providing information about productivity of corresponding authors between 2014 - 2018. More specificially, the following data points from the Web of of Science will be investigated:

- Web of Science collections SCI, SSCI and WOS.AHCI 
- Article types Original Articles and Reviews
- Country of affiliations
- ISSN
- Publication year

This dataset will be extended with:

- Publisher and journal titles from Crossref
- Journal open access status provided by the Bielefeld GOLD OA list

In this document, every step needed to create the dataset is described. This document is written in [RMarkdown](https://rmarkdown.rstudio.com/). It can be re-compiled calling

```r
require(rmarkdown)
require(tidyverse)
require(RJDBC)
require(rJava)
require(writexl)

rmarkdown::render("kb_fetch_new.Rmd")
```

@@

```{r}
# deal with rJava memory allocation 
# https://stackoverflow.com/questions/34624002/r-error-java-lang-outofmemoryerror-java-heap-space
options(java.parameters = "-Xmx1024m")
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

@@

```{sql connection=jdbcConnection, output.var="rp_country"}
select
        wos_b_2019.d_items_authors_institutions.inst_countrycode,
        d.pubyear,
        d.issn,
        count(distinct(d.ut_eid)) as pubs
    from
        wos_b_2019.d_items_authors_institutions           
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                wos_b_2019.items.pubyear,
                issn             
            from
                wos_b_2019.items                               
            inner join
                wos_b_2019.databasecollection                                                                     
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items               
            inner join
                wos_b_2019.items_authors_institutions                                     
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items             
            inner join
                wos_b_2019.issues                                   
                    on wos_b_2019.issues.pk_issues = wos_b_2019.items.fk_issues                
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                                  
                )                                                             
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'                                  
                )                                                              
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018                                  
                )                                                         
                and wos_b_2019.items_authors_institutions.type = 'RP'                           
        ) d                           
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items                                   
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors                          
            )               
    where
        wos_b_2019.d_items_authors_institutions.inst_countrycode is not null     
    group by
        wos_b_2019.d_items_authors_institutions.inst_countrycode,
        d.pubyear,
        d.issn
```

```{r}
# dump
write_csv(rp_country, "data/rp_data_14_18.csv")
```

@@

```{r}
cr_journals <- readr::read_csv("data/cr_wos.csv") %>%
  # merge springer nature brands
  mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
rp_country <- readr::read_csv("data/rp_data_14_18.csv")
issn_l <- readr::read_tsv("data/20190818.ISSN-to-ISSN-L.txt") %>%
  # manual fix Journal - American Water Works 
  add_row(ISSN = "2164-4535", `ISSN-L` = "0003-150X")
rp_df <- rp_country %>%
  left_join(issn_l, by = "ISSN") %>%
  left_join(cr_journals, by = c(`ISSN-L`= "issn_l")) %>%
  select(-issn) %>%
  distinct()
```

@@

```{r}
rp_df %>% 
  distinct(ISSN, publisher) %>% 
  count(publisher, sort = TRUE) %>% 
  mutate(prop = n /sum(n))
```

@@

```{r}
rp_df_oa %>%
  select(
    issn_wos = ISSN,
    country_code = INST_COUNTRYCODE,
    publication_year = PUBYEAR,
    articles = PUBS,
    journal_title, publisher, oa_journal,
    issn_l = `ISSN-L`) %>%
  write_csv("data/rp_jn_14_18.csv")
```

Excel dump. To avoid loading issues because of too many rows in one sheet, split the dataset into several sheets per year.

```{r}
require(writexl)
rp_df <- readr::read_csv("data/rp_jn_14_18.csv")
tmp <- map(2014:2018, function(x) {
  rp_df %>% 
    filter(publication_year == x)
  }
)
# names
names(tmp) <- paste0("pubyear_", 2014:2018)
write_xlsx(tmp, "data/rp_14_18.xlsx")
```

@@

|Variable           |Description                                                                                                                                                                      |Source                                                                                                                                                       |
|:------------------|:--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|:------------------------------------------------------------------------------------------------------------------------------------------------------------|
|`issn_wos`         |ISSN, a standardized journal id.                                                                                                                                                 |KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                  |
|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                                |
|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                                                             |KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                                |
|`articles`         |Number of original articles and reviews published. Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally. |KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`,  `wos_b_2019.d_items_authors_institutions.inst_countrycode` and  `wos_b_2019.items.pubyear` |
|`journal_title`    |Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                       |Crossref                                                                                                                                                     |
|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                      |Crossref                                                                                                                                                     |
|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?                                                                                             |Bielefeld GOLD OA List V3                                                                                                                                    |
|`issn_l`           |Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.                                             |CIEPS                                                                                                                                                        |

>>>>

---
title: "Publisher League Global"
author: "Najko Jahn"
date: "9/5/2019"
output: github_document
---

@@

This executable report describes how data about the global market shares of publishers were retrieved from the Web of Science (WoS) in-house database hosted by the [German Competence Center for Bibliometrics (KB)](http://www.bibliometrie.info/). This document is written in R Markdown. If you have access to WoS-KB data infrastructure, you will be able to replicate the methods used by calling

```r
require(rmarkdown)
require(tidyverse)
require(RJDBC)
require(rJava)
require(writexl)
rmarkdown::render("kb_fetch_publisher_global.Rmd")
```

@@

```{r}
# deal with rJava memory allocation 
# https://stackoverflow.com/questions/34624002/r-error-java-lang-outofmemoryerror-java-heap-space
options(java.parameters = "-Xmx1024m")
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar") # you may need to change the path to your db driver
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    # login credentials are stored in the .Rprofile file 
    Sys.getenv("kb_user"), 
    Sys.getenv("kb_pwd")
  ) 
```

@@

```{sql connection=jdbcConnection, output.var="rp_publisher"}
select
        wos_b_2019.items.pubyear,
        wos_b_2019.issues.issn,
        count(distinct(ut_eid)) as pubs                         
    from
        wos_b_2019.items                                            
    inner join
        wos_b_2019.databasecollection                                                                                          
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                                  
    inner join
        wos_b_2019.issues                                                        
            on wos_b_2019.issues.pk_issues = wos_b_2019.items.fk_issues                             
    where
        wos_b_2019.databasecollection.edition_value in (
            'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                                                   
        )                                                                              
        and wos_b_2019.items.doctype in (
            'Article', 'Review'                                                   
        )                                                                               
        and wos_b_2019.items.pubyear in (
            2014, 2015, 2016, 2017, 2018                                                   
        )                                                  
    group by
        wos_b_2019.items.pubyear,
        wos_b_2019.issues.issn
```

```{r}
head(rp_publisher) %>%
  knitr::kable()
# dump
write_csv(rp_publisher, "data/publisher_league_14_18.csv")
```

@@

```{r}
cr_journals <- readr::read_csv("data/cr_wos.csv") %>%
  # merge springer nature brands
  mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
rp_publisher <- readr::read_csv("data/publisher_league_14_18.csv")
issn_l <- readr::read_tsv("data/20190818.ISSN-to-ISSN-L.txt") %>%
  # manual fix Journal - American Water Works 
  add_row(ISSN = "2164-4535", `ISSN-L` = "0003-150X")
rp_df <- rp_publisher %>%
  left_join(issn_l, by = "ISSN") %>%
  left_join(cr_journals, by = c(`ISSN-L`= "issn_l")) %>%
  select(-issn) %>%
  distinct()
```

@@

```{r}
rp_df %>% 
  distinct(ISSN, publisher) %>% 
  count(publisher, sort = TRUE) %>% 
  mutate(prop = n /sum(n) * 100)
```

Next, the number and proportion of original articles and reviews published per publisher is presented:

```{r}
rp_df %>% 
  group_by(publisher) %>%
  summarise(n = sum(PUBS)) %>% 
  mutate(prop = n /sum(n) * 100) %>%
  arrange(desc(prop))
```

@@

```{r}
rp_df_oa %>%
  select(
    issn_wos = ISSN,
    publication_year = PUBYEAR,
    articles = PUBS,
    journal_title, publisher, oa_journal,
    issn_l = `ISSN-L`
    ) %>%
  write_csv("data/publisher_14_18.csv")
```

Dump to Excel

```{r}
readr::read_csv("data/publisher_14_18.csv") %>%
  writexl::write_xlsx("data/publisher_14_18.xlsx")
```

@@

|Variable           |Description                                                                                                                                                                      |Source                                                                                                                                                       |
|:------------------|:--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|:------------------------------------------------------------------------------------------------------------------------------------------------------------|
|`issn_wos`         |ISSN, a standardized journal id.                                                                                                                                                 |KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                  |
|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                                                             |KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                                |
|`articles`         |Number of original articles and reviews published. |KB Web of Science: Grouped counts over `wos_b_2019.issues.issn` and  `wos_b_2019.items.pubyear` |
|`journal_title`    |Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                       |Crossref                                                                                                                                                     |
|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                      |Crossref                                                                                                                                                     |
|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?                                                                                             |Bielefeld GOLD OA List V3                                                                                                                                    |
|`issn_l`           |Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.                                             |CIEPS                                                                                                                                                        |
